Quantitative Auditing of AI Fairness with Differentially Private Synthetic Data

### 2024-12-17 paper/paper.tex
Eval

Let us inspect the definitions of the fairness measures one by one by expanding them by the definition of conditional probability and coloring them according to whether they are in the original data or not; we will color the variables that are in the original data green and those not red.

\definecolor{lightgreen}{rgb}{0.6, 1.0, 0.6}
\definecolor{lightred}{rgb}{1.0, 0.6, 0.6}
\newcommand{\mathhighlight}[2]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\y}[1]{\mathhighlight{lightgreen}{#1}}
\newcommand{\n}[1]{\mathhighlight{lightred}{#1}}

@@

Now we can count the numbers of available variables in the original data for each fairness measure: Demographic Parity has $\frac{6}{8} = 0.75$; Overall Accuracy Equality has $\frac{4}{6} = 0.6\overline{6}$; Equalized Odds has $\frac{8}{10} = 0.8$; and Conditional Use Accuracy Equality has $\frac{6}{10} = 0.6$.

Hence, we can say less about fairness with the measures with less information, and the measures approximated by synthetic data may be less accurate.

Now, looking at the actual experimental results, first and foremost, across all datasets, we observed that Conditional Use Accuracy Equality did indeed show higher errors than other measures. For example, in the COMPAS experiment, both Conditional Use Accuracy Equality measures had an error over 0.1. This corroborates our assumption that it is harder to approximate sufficiency measures on synthetic data.

This is a fundamental limitation of this approach, and it is important to be aware of this when interpreting the results of fairness analyses. The auditor needs to keep in mind that they are approximating scores, and the scores may not be fully accurate. The auditor should be aware that they are approximating fairness.

@@

@@

>>>>

Let us inspect the definitions of the fairness measures one by one by expanding them by the definition of conditional probability and marking them according to whether they are in the original data or not; we will box the variables that are not in the original data.

\definecolor{lightgreen}{rgb}{0.6, 1.0, 0.6}
\definecolor{lightred}{rgb}{1.0, 0.6, 0.6}
\newcommand{\mathhighlight}[2]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\y}[1]{{#1}}
\newcommand{\n}[1]{\boxed{#1}}

@@

Here we observed that, of the measures considered, only Conditional Use Accuracy Equality has boxed variables that are in both the numerator and the denominator. This is in contrast to the other measures, which only have one boxed variable in the numerator.

We can say less about fairness with the measures with less information, and thus the measures approximated by synthetic data may be less accurate for Conditional Use Accuracy Equality.

@@

In addition, further experiments showed that, in an alternative assumption where the auditor has knowledge of $(S,Y)$ and can add the edge $(S,Y)$ to the Markov random field before fitting, the synthesized data would have been more accurate, even lowering the error of Conditional Use Accuracy Equality in the COMPAS example.

@@

Auditors hence gained an approach to auditing AI fairness alternative to the traditional method of using real data. When using synthetic data derived from real data, the fairness properties remain meaningful and can still be enough for auditing purposes.

Looking at the actual experimental results, first and foremost, across all datasets, we observed that Conditional Use Accuracy Equality did indeed show higher errors than other measures.

In the Adult experiment, Conditional Use Accuracy Equality (True Positive) has an error of around 0.16. In the COMPAS experiment, both Conditional Use Accuracy Equality measures had an error over 0.1. In the Diabetes experiment, Conditional Use Accuracy Equality (True Negative) has an error much higher than other measures.

This corroborates our assumption that it is harder to approximate sufficiency measures on synthetic data. As there is less information available for the synthesis process, the measures that require more information will be harder to approximate.

Although further experiments showed that linking $(S,Y)$ before fitting can somewhat alleviate this problem, the auditor may not always have the background knowledge to identify them; moreover, this improvement is not always guaranteed.

This is a fundamental limitation of this approach, and it is important to be aware of this when interpreting the results of fairness analyses. The auditor needs to keep in mind that they are approximating scores, and the scores may not be fully accurate.
